Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-14 main.tex
Notes
Thoughts

@@

Formally, a policy
%
$\pol \in \CONTOURS \rightarrow \exists X.\; X \times (\MACHS \times X
\rightharpoonup \MACHS \times X)$
%
 is a function that takes a contour to
\begin{enumerate*}
  \item a private type $X$,
  \item an initial value of $X$, and
  \item a partial step function of type
  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
  agrees with $\stepstounderfull$ on machine states.
\end{enumerate*}
That is, a policy takes a starting contour and produces a runtime monitor
that carries along some private information and uses it to decide whether to
terminate execution early.

@@

\bcp{Here is where it starts getting iffy...}

A {\em contoured (hyper)property} is a function that takes a contour to a
hyperproperty.  That is, it takes a starting contour and produces a
\bcp{prefix-closed?} set of pairs of sequences of machine states.

The {\em observation trace} arising from a finite or infinite sequence of
contours and machine states
$(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$, written
$\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$ is a
partial map taking each machine component to the (finite or infinite)
sequence of {\em distinct} values taken on by this component at moments when
it is {\em observable}.  Formally, this map can be constructed as
follows. For each component $\component$,
%
(1) if
$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
$\obstrace(\component) = \bot$;
%
and otherwise (2), $\obstrace(\component)$ is a sequence whose first element
is $\mach_i(\component)$ for the smallest $i$ such that
$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
than the index $i$ of the $m^{th}$ element and such that
$\contour_j(\component) = (\underscore,\OBS)$ and
$\mach_j(\component) \neq \mach_i(\component)$.
%
\iftext
\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
\newcommand{\aux}{\mathit{aux}}
\bcp{Don't like the ``rest'' argument and the fact that we don't
  have a metavariable for sequences of pairs of contours and machines.
  Consider defining one.}
%
\[
\begin{array}{lcl}
\aux \; ( [\,] ) \; \trace& = & \trace
\\
\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
  & = &
        \aux \;( \mathit{rest} )
        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
\\
\\
\obstrace \;( (\contour_0,\mach_0),\, \ldots )
  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
      \component.\, \bot)
\end{array}
\]
\fi

@@

>>>>

The {\em observation trace} arising from a finite or infinite sequence of
pairs of a contour and a machine state
$(\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots$, written
$\obstrace((\contour_0,\mach_0),\, (\contour_1,\mach_1),\ldots)$ is a
partial map taking each machine component to the (finite or infinite)
sequence of the {\em distinct} values taken on by this component at moments
when it is {\em observable}.  Formally, this map can be constructed as
follows. For each component $\component$,
%
(1) if
$\contour_i(\component) = (\underscore,\HIDDEN)$ for all $i$, then
$\obstrace(\component) = \bot$;
%
and otherwise (2), $\obstrace(\component)$ is a sequence whose first element
is $\mach_i(\component)$ for the smallest $i$ such that
$\contour_i(\component) = (\underscore,\OBS)$ and whose $(m+1)^{th}$ element
(if any) is $\mach_j(\component)$, where $j$ is the smallest index greater
than the index $i$ of the $m^{th}$ element and such that
$\contour_j(\component) = (\underscore,\OBS)$ and
$\mach_j(\component) \neq \mach_i(\component)$.
%
\iftext
\bcp{Unfinished and horrible...  But maybe the textual description is OK?}
\newcommand{\aux}{\mathit{aux}}
\bcp{Don't like the ``rest'' argument and the fact that we don't
  have a metavariable for sequences of pairs of contours and machines.
  Consider defining one.}
%
\[
\begin{array}{lcl}
\aux \; ( [\,] ) \; \trace& = & \trace
\\
\aux \; ( (\contour,\mach) \mathrel{::} \mathit{rest} ) \; T
  & = &
        \aux \;( \mathit{rest} )
        \\ && \quad (\lambda k.\; \mbox{\bcp{and so on}}
\\
\\
\obstrace \;( (\contour_0,\mach_0),\, \ldots )
  &=& \aux \; ( (\contour_0,\mach_0),\, \ldots ) \; (\lambda
      \component.\, \bot)
\end{array}
\]
\fi

@@

\bcp{Fix: even simpler and cleaner, just run the step function to completion
  and forget about $X$---i.e., a property is just a function from a pair of
  a contour and a machine state to a finite or infinite sequence of machine
  states!}  Formally, a policy
%
$\pol \in \exists X.\; (\CONTOURS \rightarrow X) \times (\MACHS \times X
\rightharpoonup \MACHS \times X)$
%
comprises
\begin{enumerate*}
  \item a private type $X$,
  \item a function taking a contour to an initial value of $X$, and
  \item a partial step function of type
  $(\MACHS \times X) \rightharpoonup (\MACHS \times X)$ that, when defined,
  agrees with $\stepstounderfull$ on machine states. \bcp{We could actually
    just make it return an X and calculate the machine state from the
    ordinary step function.  This is perhaps a bit cleaner.}
\end{enumerate*}
That is, a policy takes a starting contour and produces a runtime monitor
that carries along some private information and uses it to decide whether to
terminate execution early.

@@

% A {\em contoured (hyper)property} is a function that takes a contour to a
% hyperproperty.  That is, it takes a starting contour and produces a
% \bcp{prefix-closed?} set of pairs of sequences of machine states.

\bcp{It gets sketchy from here on!  The observer needs to be simplified as
  above, at a minimum.}

An {\em observer} $\observer \in \exists X.\; (\CONTOURS \rightarrow X) \times
(X \rightarrow \CONTOURS) \times (\MACHS \times X \rightharpoonup \MACHS
\times X) $
comprises
\begin{enumerate*}
  \item a private type $X$,
  \item a function taking a contour to an initial value of $X$, and
  \item a function that allows ``reading off'' a contour from an $X$, and
  \item a partial step function of type
  $(\MACHS \times X) \rightharpoonup X$.
\end{enumerate*}

@@

  \bcp{Answer: Just make instructions read-only.}
